### 2024-08-26 manuscript.tex
Multiple updates

Together, we analysed data from 31 voyages of RV \emph{Polarstern}, resupply
vessel (RSV) \emph{Aurora Australis}, RV \emph{Tangaroa}, RV \emph{Nathaniel B.
Palmer}, Her Majesty's New Zealand Ship (HMNZS) \emph{Wellington} and one
sub-Antarctic station (Macquarie Island) in the SO south of 40°S between year
2010 and 2021 (Fig. \ref{fig:map}, Table \ref{tab:voyages} and
\ref{tab:voyage-references}). Altogether, the voyages and the station comprised
2421 days of data south of 40°S, but the availability of ceilometer data was
smaller.

Missing days in the ceilometer data were HMNZSW16 (7 days): 24--27 November, 10
December, 16--17 December 2016; MARCUS (3 days): 8, 10 November, 10 December;
MICRE (9 days): 7--8, 29 June, 5, 16 July, 15 August, 17 October 2016, 11
February, 21 March 2017; TAN1502 (1 day): 24 January.

These data sources contained ceilometer (lidar) observations with the Vaisala
CL51 and CL31 operating at 910 nm and the Lufft CHM-15k operating at 1064 nm. A
ceilometer is a low-power vertically pointing lidar with the purpose of
measuring the cloud base, but it also measures the full vertical structure of
clouds as long as the laser signal is not attenuated by thick clouds. A total
of about 1500 days of lidar observations were included in our analysis.

Apart from lidar observations, radiosondes were launched on weather balloons
at regular synoptic times on the RV \emph{Polarstern} voyages, and surface
meteorological quantities were measured continuously.

@@

\subsection{Vaisala CL51 and CL31}

@@

\subsection{Lufft CHM 15k}

@@

It was not possible to include precipitation simulation in the ALCF due to the
absence of required fields in our model output and reanalysis data (liquid and
ice precipitation mass mixing ratios) because the required radiation
calculations are not implemented in the ALCF. At the same time, it is relatively
difficult to distinguish precipitation backscatter from cloud backscatter
in lidar observations, especially when only one wavelength channel and no
polarisation channel are available.

In order to get a fair comparison of observations and models, one needs to
exclude profiles with precipitation either manually or using an automated
method. In models, this can be accomplished relatively easily by excluding
profiles with an amount of surface precipitation flux exceeding a threshold.
In observations, this can be more complicated, especially on voyages, because
rain gauges are known to be unreliable on ships.

On the RV \emph{Polarstern} voyages, regular synoptic observations were
available.  These included precipitation observations by an observer. We
used this dataset to train a convolutional artiﬁcial neural network (ANN)
of the U-Net type \citep{ronneberger2015} to recognise precipitation in the
lidar data. Samples of short time intervals (10 min) of near-surface
backscatter (0–250 m) were classified as clear, rain, snow and fog, using the
synoptic observations as a training set. From these, a binary classification of
profiles as wet or dry was derived.

@@

\begin{figure}
\centering
\includegraphics[width=\textwidth]{img/cl_agg_exp.pdf}
\caption{
Cloud occurrence by height calculated as the average of all voyages and stations
for the observed (OBS) and simulated lidar cloud mask. The total cloud fraction
(CF), average shortwave (SW) and longwave (LW) are shown in the legend. The
bands span the 16$^\mathrm{th}$--84$^\mathrm{th}$ percentile calculated from
the set of all voyages and stations. The subsets \textbf{(a--g)} are described
in the text (Section \ref{sec:results}).}
\label{fig:cloud-occurrence}
\end{figure}

@@

We compared about 2000 radiosonde proﬁles from 24 RV \emph{Polarstern}
voyages between the observations and the model. Proﬁles in the model were taken
at the same geographical location and time relative to the start of the year.
Only proﬁles for which the sea ice conditions (sea ice present or absent) are
the same in the observations and the model were included. We derived potential
temperature profiles, lifting condensation level (LCL), and aggregated the
profiles by sea ice conditions (Fig. \ref{fig:radiosonde}).

>>>>

Together, we analysed data from 31 voyages of RV \emph{Polarstern}, resupply
vessel (RSV) \emph{Aurora Australis}, RV \emph{Tangaroa}, RV \emph{Nathaniel B.
Palmer}, Her Majesty's New Zealand Ship (HMNZS) \emph{Wellington} and one
sub-Antarctic station (Macquarie Island) in the SO south of 40°S between year
2010 and 2021 (Fig. \ref{fig:map}, Table \ref{tab:voyages} and
\ref{tab:voyage-references}). Altogether, the voyages and the station comprised
2421 days of data south of 40°S, but the availability of ceilometer data was
smaller due to gaps in measurements. Missing days in the ceilometer data were
HMNZSW16 (7 days): 24--27 November, 10 December, 16--17 December 2016; MARCUS
(3 days): 8, 10 November, 10 December; MICRE (9 days): 7--8, 29 June, 5, 16
July, 15 August, 17 October 2016, 11 February, 21 March 2017; TAN1502 (1 day):
24 January.

These data sources contained ceilometer observations with the Vaisala CL51 and
CL31 operating at 910 nm and the Lufft CHM-15k operating at 1064 nm, described
in detail below (Sections \ref{sec:cl51} and \ref{sec:chm15k}). A ceilometer is
a low-power near-infrared vertically pointing lidar with the purpose of
measuring the cloud base, but it also measures the full vertical structure of
clouds as long as the laser signal is not attenuated by thick clouds.

Apart from lidar observations, radiosondes were launched on weather balloons at
regular synoptic times on the RV \emph{Polarstern}, MARCUS, NBP17024, TAN1702,
and TAN1802 voyage, measuring pressure, temperature, relative humidity and
global navigation satellite system coordinates. Derived thermodynamic (virtual
potential temperature, lifting condensation level, \ldots) and dynamic
quantities (wind speed and direction) for the measured vertical profiles were
calculated with rstool \citep{rstool}. Surface meteorological quantities were
measured continuously by an onboard automatic weather station or individual
instruments.

@@

\subsection{Vaisala CL51 and CL31}
\label{sec:cl51}

@@

\subsection{Lufft CHM 15k}
\label{sec:chm15k}

@@

Precipitation can cause strong enough lidar backscattering to be recognised as
a cloud by the cloud detection method used in the ALCF. This is undesirable
if equivalent precipitation backscatter is not included in the simulated lidar
profiles. It was not possible to include precipitation simulation in the ALCF
due to the absence of required fields in the ICON model output and the
reanalysis data, the liquid and ice precipitation mass mixing ratios, and
because the required radiation calculations for precipitation are not
implemented in the ALCF. In this situation, in order to achieve a fair
comparison of observations and models, one should exclude profiles with
precipitation either manually or using an automated method. It is relatively
difficult to distinguish precipitation backscatter from cloud backscatter in
lidar observations, especially when only one wavelength channel and no
polarisation channel are available. In models, the same can be accomplished
relatively easily by excluding profiles with an amount of surface precipitation
flux exceeding a given threshold. In the observations, using precipitation
flux measurements can be very unreliable on ships due to ship movement and
turbulence caused by nearby ship structures. Our analysis of rain gauge data
from the RV \emph{Tangaroa} showed large discrepancies between the rain gauge
time series and human-performed synoptic observations, as well as large
inconsistencies in the rain gauge time series. Therefore, it is desirable to
implement a method of detecting precipitation from observed backscatter
profiles instead.

On the RV \emph{Polarstern} voyages, regular human-performed synoptic
observations were available and included precipitation presence and type. We
used this dataset to train a convolutional artiﬁcial neural network (ANN) of
the U-Net type \citep{ronneberger2015} to recognise profiles with precipitation
from lidar backscatter. Samples of short time intervals (10 min) of
near-surface backscatter (0–250 m) were classified as clear, rain, snow and
fog, using the synoptic observations as a training dataset. From these, a
binary classification of profiles as wet (rain or fog) and dry (clear or fog)
was derived.

@@

\begin{figure}
\centering
\includegraphics[width=\textwidth]{img/cl_agg.pdf}
\caption{
Cloud occurrence by height calculated as the average of all voyages and stations
for the observed (OBS) and simulated lidar cloud mask. The total cloud fraction
(CF), average shortwave (SW) and longwave (LW) are shown in the legend. The
bands span the 16$^\mathrm{th}$--84$^\mathrm{th}$ percentile calculated from
the set of all voyages and stations. The subsets \textbf{(a--g)} are described
in the text (Section \ref{sec:results}).}
\label{fig:cloud-occurrence}
\end{figure}

@@

We analysed about 2300 radiosonde proﬁles south of 40°S from the 24 RV
\emph{Polarstern} voyages, MARCUS, NBP1704, TAN1702, and TAN1802.

voyages between the observations and the model. Proﬁles in the model were taken
at the same geographical location and time relative to the start of the year.
Only proﬁles for which the sea ice conditions (sea ice present or absent) are
the same in the observations and the model were included. We derived potential
temperature profiles, lifting condensation level (LCL), and aggregated the
profiles by sea ice conditions (Fig. \ref{fig:radiosonde}).
